Centralized Programming for Distributed Control Planes

### 2016-01-27 paper/introduction.tex

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common configuration patterns are captured as parameterized templates.
 While templates help ensure certain kinds of consistency across devices, they do not
bridge the
semantic divide between network policies and device-level configuration.
They do not provide fundamentally different abstractions
from existing configuration languages.

@@

The centralized control planes of SDN, however, are not a panacea.
%
They require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of larger networks and to react sufficiently quickly
to a changing environment. To address this challenge, some researchers are beginning to 
explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.  However, academic
language design and implementation efforts have not kept pace.  For instance, work on
experimental SDN languages
such as Frenetic~\cite{frenetic}, Flowlog~\cite{flowlog}, Vericon~\cite{vericon}, Merlin~\cite{merlin}, NetKAT~\cite{netkat}, and Kinetic~\cite{kinetic}, to
name a few, has not yet shown how to implement fault tolerant
multi-controller systems that support their high-level abstractions efficiently.

In this paper, we ask if it is possible to program 
distributed control planes using network-wide policies.  In particular,
operators should be able to directly express \emph{end-to-end}, 
\emph{network-wide} policies directly; it is the job of a compiler
to decompose such policies into individual device configurations. 
In addition, the resulting forwarding behavior, which emerges through device interactions, must be policy-compliant under all possible failures.
If successful, this approach would combine easy programmability of centralized control planes with the failure robustness and scalability of distributed control planes.
%
More pragmatically, it would help many networks that, for the foreseeable future, will continue to use a distributed control plane, due to the difficulty of migrating to SDN or the inherent scalability and failure-robustness of distributed control.

Through \sysname, our system that includes a language and a compiler, we demonstrate the feasibility of programming distributed control planes. We focus on BGP, a common and highly flexible way to implement distributed control planes, and show how to automatically generate router BGP configurations from network-wide policies.

We face two challenges in designing \sysname that differ from designing a system for programming data planes~\cite{x,y,z}. The first is policy specification itself---specifying network {\em policies} is different from specifying network {\em paths}.
%
The specification must compactly capture behavior under all possible failures. Since there is no controller at runtime, the routers must be programmed ahead of time, based on the specification, to handle failures.
%
Further, many policies naturally under-specify paths (e.g., "prefer paying neighbors" is not a concrete path).
%
\sysname addresses this specification challenge by allowing operators to express {\em path preferences}, where preference describes a set of valid paths and less-preferred options are taken only when a higher-preference options are unavailable.

The direct nature of \sysname specifications brings up the second challenge, of compiling them to router configurations.  We must compute the sets of paths represented by the intersection of multiple preferences and topology, compute which ones can be honored under a given failure scenario, and ensure policy compliance under all possible failure cases. We handle this challenge by compactly capturing policy and topology in a {\em product graph} and developing efficient algorithms that operate on this graph.

We evaluate \sysname by using it to encode policies of real backbone and data center networks.

\todo{new intro is not complete}

>>>>

To reduce configuration errors, many practitioners have a adopted a template-based
approach~\cite{hatch,thwack}, in which common configuration patterns are captured as parameterized templates.  More powerful still are systems like
ConfigAssure~\cite{narain:lisa05,narain+:configassure}, which use SAT solving
and model finding tools to fill in parameters in configurations while
ensuring key correctness constraints are satisfied.  However,
 while templates help ensure certain kinds of consistency across devices, 
they do not bridge the
semantic divide between network policies and device-level configuration,
nor do they do not provide fundamentally different abstractions
from existing configuration languages or new, higher-level reasoning 
principles.

@@

The centralized control planes of SDN, however, are not a panacea.
%
They require careful design and engineering to be robust to failures at scale---one must ensure that all devices can communicate with the controller at all times, even under arbitrary combinations of failure. Even ignoring failures, it is necessary for controller systems to
scale to meet the demands of larger, geographically distributed networks, 
and to react sufficiently quickly
to a changing environment. To address this challenge, some researchers are beginning to 
explore
multi-controller networks, with interacting controllers, thus bringing back distributed
control planes~\cite{mccauley2013extending,onos} and their current programming difficulties.  
%However, academic
%language design and implementation efforts have not kept pace.  For instance, work on many
%experimental SDN languages~\cite{frenetic,flowlog,vericon,merlin,netkat,kinetic,pga} has not yet shown how to implement fault tolerant
%multi-controller systems that support their high-level abstractions efficiently.
A second problem with existing SDN frameworks, and the languages used to
program them, is that while they provide many 
effective \emph{intra}-domain routing
abstractions, they fail to provide a \emph{coherent} means to enable 
\emph{inter}-domain communication of available routes.

Hence, in this paper, we have two central goals:
\begin{enumerate}
\item Design a new, high-level, declarative language with abstractions 
suitable for simultaneously expressing intra-domain routing, inter-domain 
routing and alternatives in case of failures.
\item Define algorithms for compiling these specifications in to 
configurations for off-the-shelf devices running standard
distributed control plane algorithms, while ensuring correct behavior
independent of the number of faults.
\end{enumerate}

To achieve the first goal, we borrowed heavily from the design of
recent high-level SDN languages based around specification of 
paths via regular expressions such as FatTire~\cite{fattire}, 
Merlin~\cite{foster:merlin}, and
NetKAT~\cite{netkat}.  However, one of the key differences is that
we incorporate expression of user preferences in to the algebra
of regular expressions and provide a semantics to our language
in terms of sets of ranked paths.  The preferences indicate possible
fail-over routes:  If too many faults occur, the network may be
disconnected, but as long as any path the user has expressed is
still available, the system guarantees the highest available ranked
path will be taken.  We also support network operators by defining a
set of common abbreviations for entering, leaving, and traversing
the user network.  Operators use familiar logical operations
to build routes in a modular and compositional manner from these 
primitives.  All abbreviations are translated in to our regular-expression-based
intermediate representation and compiled from there.

To achieve the second goal, we define program analysis and compilation
algorithms that translate the regular policies in to a graph-based
intermediate language and from there in to BGP.  During the
compilation process, our compiler automatically synthesizes per-device
import and export filters, local preferences, MED attributes, and 
community tags.  We chose BGP for pragmatic reasons:  it is an industry
standard and many data centers use it internally as well as externally.
Indeed, despite the presence of SDN, many networks will continue to
use BGP for the foreseeable future due to investments they have already
made in the infrastructure, the difficulty of transitioning to SDN,
and the scalability and fault-tolerance advantages of a distributed
control plane.  However, using BGP as a back end is not without its 
challenges: not all high-level specifications can be implemented correctly in
the presence of arbitrary failures.  Consequently, we define new
program analysis algorithms to detect specifications that cannot be implemented
and report our findings to the user, so he or she may fix the specification
early rather than waiting until configurations have been deployed.  
These algorithms differ from
standard BGP analysis algorithms~\cite{feamster:rcc,batfish} because they operate on representations of
our high-level specifications (which has the pleasant effect of simplifying our 
task) rather than directly on arbitrary BGP configurations.

To evaluate language and system, we have experimented with specifying policy
for several realistic data center and backbone networks. We have also
measured the performance of our compiler, demonstrating that our compilation
algorithms scale to topologies with several thousand routers.

In the remainder of the paper, we review BGP~\ref{sec:background} (readers
familiar with BGP may skip this section),
further motivate the need for new languages 
for programming distributed control planes (Section \ref{sec:motivation}),
introduce \sysname{} by example (Section \ref{sec:propane}), 
explain our compiler algorithms (Section \ref{sec:compilation}),
present several lemmas justifying the correctness of our techniques (Section \ref{sec:theory}), describe our implementation (Section \ref{sec:implementation}
and discuss related work (Section \ref{sec:related}).

%% In this paper, we ask if it is possible to program 
%% fully distributed control planes using 
%% \emph{end-to-end}, 
%% \emph{network-wide} policies, while handing the job of compiling
%% such policy.  In particular,
%% operators should be able to directly express \emph{end-to-end}, 
%% \emph{network-wide} policies directly; it is the job of a compiler
%% to decompose such policies into individual device configurations. 
%% In addition, the resulting forwarding behavior, which emerges through device interactions, must be policy-compliant under all possible failures.
%% If successful, this approach would combine easy programmability of centralized control planes with the failure robustness and scalability of distributed control planes.
%% %
%% More pragmatically, it would help many networks that, for the foreseeable future, will continue to use a distributed control plane, due to the difficulty of migrating to SDN or the inherent scalability and failure-robustness of distributed control.

%% Through \sysname, our system that includes a language and a compiler, we demonstrate the feasibility of programming distributed control planes. We focus on BGP, a common and highly flexible way to implement distributed control planes, and show how to automatically generate router BGP configurations from network-wide policies.

%% We face two challenges in designing \sysname that differ from designing a system for programming data planes~\cite{x,y,z}. The first is policy specification itself---specifying network {\em policies} is different from specifying network {\em paths}.
%% %
%% The specification must compactly capture behavior under all possible failures. Since there is no controller at runtime, the routers must be programmed ahead of time, based on the specification, to handle failures.
%% %
%% Further, many policies naturally under-specify paths (e.g., "prefer paying neighbors" is not a concrete path).
%% %
%% \sysname addresses this specification challenge by allowing operators to express {\em path preferences}, where preference describes a set of valid paths and less-preferred options are taken only when a higher-preference options are unavailable.

%% The direct nature of \sysname specifications brings up the second challenge, of compiling them to router configurations.  We must compute the sets of paths represented by the intersection of multiple preferences and topology, compute which ones can be honored under a given failure scenario, and ensure policy compliance under all possible failure cases. We handle this challenge by compactly capturing policy and topology in a {\em product graph} and developing efficient algorithms that operate on this graph.

%% We evaluate \sysname by using it to encode policies of real backbone and data center networks.

%% \todo{new intro is not complete}
